---
title: "Dynamic Landmarking on SPLC Data"
author: "Anya Fries"
date: "17/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data loading and preparation

*Note: hist_final is one-hot encoded, with SC used as a reference, in order to allow for the operation: coefficient $\times$ LM*

#### Load data and libraries

```{r libs, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(survival)
library(dynpred)
library(riskRegression)
library(msm) # for delta method
library(knitr) # for kable to print tables

source("Anya SPLC-RAT Dev Data/source_general.R")
source("dyn-landmarking.R")
source("SPLC-helper-functions.R")

MYDAT <- read.csv("Anya SPLC-RAT Dev Data/DAT_splcrat.csv") 
MYDAT <- MYDAT[MYDAT$Time>0,]
```


#### Set up base data with relevant covariates

```{r data-clean}
covars <- c("ID_secLung","stage2.ix","hist_final","ph","USPSTF2","cigday2","surgery1")

add_vars <- c("packyears2",  "smkstatus2", "quityears2","age_ix",
              "packyears_fup","smkstatus_fup","cigday_fup",
              "diff_fup_IPLC") # Time from FUP to IPLC
all_vars <- c(covars,add_vars)

MEC <- MYDAT %>% 
  select(Time,event,all_of(all_vars)) %>%
  mutate(USPSTF.stage = USPSTF2 * stage2.ix,
         value = 1)  %>% 
  spread(hist_final, value,  fill = 0 ) %>% 
  select(-SC) # one-hot encodings of hist_final

hist_final_labels <- c("AD","LC","NSCLC_NOS","OTH","SQ")
covars <- c(covars[-3], "USPSTF.stage", hist_final_labels) 
all_vars <- c(covars,add_vars)

 # reorder to have add_vars at the end
MEC <- MEC %>% select(Time,event,ID_secLung:surgery1,USPSTF.stage:SQ,all_of(add_vars))
colnames(MEC)
```

```{r distr-events, out.width='60%', include=FALSE, eval=FALSE}
boxplot(MEC$Time ~ MEC$event)
```

## Create landmarking super data set

We set our specific values and build the super LM dataset. We will produce 5-year dynamic predictions of SPLC. Landmark time points were established every 6 months between 0 and 3 years.

```{r set-up-LM-interactions}
w <- 5        # risk prediction window (risk within time w)

LMs <- seq(0,3,by=0.5)
outcome <- list(time="Time",status="event")
covs <- list(fixed=all_vars,varying=NULL)

# f gives covariate-time interactions
f1 <- function(t) 1
f2 <- function(t) t
f3 <- function(t) t^2

# g lets the hazard depend on time
g1 <- function(t) f2(t)
g2 <- function(t) f3(t)

func_covars <- list(c(f1,f2,f3))
func_LMs <- list(c(g1,g2))

# Choose variables that will have time interaction
pred_covars <- covars[2:length(covars)] 
```


```{r super-data-set}
# Stack datasets
LMdata <- cutLMsuper(MEC, outcome, LMs, w, covs, format="wide")
# Update LM-varying covariates (user created function update)
LMdata <- update(LMdata)
# Add time interactions
LMdata <- addLMtime(LMdata, pred_covars, func_covars, func_LMs)

str(LMdata)
```

## Build the predictive model


Fit the supermodel using all of the covariates on the super dataset. Note that `LMdata$allLMcovars` gives all the covariates with LM interaction (`LMdata$LMcovars` are these same covariates but without interaction). A subset of these variables can be used.

```{r}
model_covars <- LMdata$allLMcovars
print(model_covars)
```


```{r fit-super-model}
source("dyn-landmarking.R")
source("SPLC-helper-functions.R")
formula <- paste("Hist(Time, event, LM)", "~", 
                 paste(model_covars, collapse=" + "), "+ cluster(ID_secLung)")
LMsupercsc2 <- fitLM("CSC", as.formula(formula), LMdata, cause=1, method="breslow") 
# LMsupercsc2
```



## Exploring the predictive model

### Exploring the coefficients of the model

```{r load-html-table, warning=FALSE}
library(htmlTable)
n_dec <- 2
get_coef_info(LMsupercsc2$superfm, n_dec, CI=FALSE, pVal=TRUE)
```


### Visualising the changing hazard ratios for different covariates

```{r plot-dyn-HR}
plot_dynamic_HR(LMsupercsc2, end_time=3, CI=TRUE)
```


## Prediction

A prediction is made for an individual at a specific prediction time (e.g., at baseline, at 2 years, etc). Thus both an individual and a prediction (landmark) time must be given. The $w$-year risk is predicted ($w$ is set when fitting the model).  

### Create newdata

New data is a dataframe with individual covariates associated with a LMing time point (i.e. covariates updated for the LMing time-point). Only the covariates used in the model need to be provided. E.g., this can be an entry from the very original data frame. Note that cutLM/cutLMsuper can be used to create these individuals for later prediction time points. 

New data can be entire dataframes associated with one LM-time, for e.g., as follows: 

```{r newdata}
times1 <- c(0,1,2)

# update time varying covariates for each of these prediction points 
# use a custom function cutLMnewdata which is the same at cutLM but without censoring time at prediction window
newdata1 <- lapply(times1, 
                  function(t){ 
                    update(cutLMnewdata(MEC, outcome, t, t+w, covs, format="wide"))
                    }
                  ) 

print("Data on which to perform 5-year risk prediction at 1 year:")
head(newdata1[[2]])
```

New data can be whole dataframes associated with multiple times, where one LM/prediction time is given for each row. Here we also make predictions at 0,1,2 years but store all the data together.

```{r}
option2 <- cutLMsuper(MEC, outcome, times1, w, covs, "wide")
newdata2 <- option2$LMdata
times2 <- newdata2$LM

head(newdata2)
head(times2)
```



#### Get predictions for each individual

Predictions can be made on either form: 

```{r get-ind-preds-all, eval=F}
risks1 <- list()

for (t in 1:length(times1)){
  tLM <- times1[t]
  newdat <- newdata1[[t]][,]
  
  risks <- predLMrisk(LMsupercsc2, newdat, tLM, cause=1)
  risks1[[t]] <- risks
}
```

```{r get-preds-all}
risks2 <- predLMrisk(LMsupercsc2, newdata2, times2, cause=1)
```


### Calibration: quantile-based calibration plot

This function builds on `calPlot` from the `pec` library. The object from `predLMrisk` already stores the time at which the prediction was made and the prediction window, so these do not need to be input. A named list of prediction objects from `predLMrisk` allows for comparison between models.  

```{r lib-pec, warning=F, message=F}
library(pec) 
```

```{r quantile-calib}
source("dyn-landmarking.R")
method="quantile"
q=10
YLIM=XLIM=c(0,0.2)

# risks2 <- risks1
# risks2$preds$risk = risks2$preds$risk + 0.03

par(mfrow=c(1,3),pty="s")
outlist = LMcalPlot(list("LM"=risks2), cause=1, method=method, q=q ,ylim=YLIM, xlim=XLIM) #,"LM2"=risks2

```


### AUCt and Brier Score 
Similarly, time-dependent AUC and Brier Score can be computed. 

Extract in a data frame:
```{r data-tab, message=F,warning=F}
library(data.table)
```

```{r LMScore, warning=F}
l2 = LMScore(list("LM1"=risks2), cause=1)
print(l2)
```


```{r score-plot, include=F, eval=F}
# Extract with calibration plot:
# TODO: add this functionality

par(mfrow=c(1,3)) 
metrics <- list()
for(t in 1:length(times)){
  tLM <- times[t]
  data_to_test <- newdata[[t]]
  risks_to_test <- preds[[t]]$Fw

  score_t = Score(list("LM"=risks_to_test), formula=Hist(Time,event,LM)~1,
              data=data_to_test, cause=1, times=c(tLM+w), plots="cal")
  plotCalibration(score_t,ylim=c(0,0.5),xlim=c(0,0.5), cens.method="local")

}
```

### Individual risk score trajectories

Individual risk score trajectories can be plotted. As with `predLMrisk`, the data input is in the form of the original data. For example:

```{r choose-inds}
idx <- c(5, 383, 880, 1044)
kable(MEC[idx,])
```

```{r, include=F, eval=F}
# # Finding interesting individuals
# MEC[!is.na(MEC$diff_fup_IPLC) &MEC$diff_fup_IPLC < 0 & MEC$smkstatus2 != MEC$smkstatus_fup &MEC$diff_fup_IPLC > -6 & MEC$Time > 6,]
```


We turn our data into long-form data to plot: (wide-form can be used too)

```{r prep-long-data}
# Prediction time points
x = seq(0,9,by=0.02)

# Create long-form data to plot individual trajetory
dat = cutLMsuper(MEC[idx,],outcome, x, w, covs, format="wide")
dat = update(dat)

head(dat$LMdata)
```


```{r plotRisk}
plotRisk(LMsupercsc2, dat$LMdata,
            format="long", LM_col = "LM", id_col="ID_secLung", cause=1,ylim=c(0,0.8))

```




